---
title: "The Laplace and two-piece Laplace Distributions"
author: '[F. Javier Rubio](https://sites.google.com/site/fjavierrubio67/)'
date: "02/07/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Laplace and two-piece Laplace Distributions

The following R code contains the implementation of the [Laplace](https://en.wikipedia.org/wiki/Laplace_distribution) and two piece Laplace distributions (using the [R package twopiece](https://sites.google.com/site/fjavierrubio67/resources)).

**References.**

1. [Inference in Two-Piece Location-Scale Models with Jeffreys Priors](http://rpubs.com/FJRubio/twopiece)

2. [Inference for grouped data with a truncated skew-Laplace distribution](https://doi.org/10.1016/j.csda.2011.06.002)

3. [twopiece R package](http://rpubs.com/FJRubio/twopiece)

4. [The Laplace Distribution and Generalizations](http://www.springer.com/gb/book/9780817641665)

```{r}
# Package only required for the last example
library(twopiece)

# Probability Density Function: location parameter mu and scale parameter sigma
dlap = function(x,mu=0,sigma=1,log=FALSE){
  log.den = - log(2) - log(sigma) - abs(x-mu)/sigma
  if(log) log.den
  else exp(log.den)
}

# Cumulative Distribution Function
plap = function(x,mu=0,sigma=1,log.p=FALSE){
    if(x<mu){
    log.cdf1 = -log(2) + (x-mu)/sigma
    if(log.p) return(log.cdf1)
    else return(exp(log.cdf1))
  }
  if(x>=mu){
      log.cdf2 = -log(2) + (-x-mu)/sigma
      if(log.p) log(1 - exp(log.cdf2))
      else 1 - exp(log.cdf2)
  }
}

# Quantile Function
qlap = function(p,mu=0,sigma=1){
  if(p<0.5) return( mu + sigma*log(2*p))
  if(p>=0.5) return( mu - sigma*log(2*(1-p)))
}

# Random number generation 
rlap = function(n,mu=0,sigma=1){
  u = runif(n)
  rgen = ifelse(u<0.5,mu + sigma*log(2*u),mu - sigma*log(2*(1-u)))
  return(rgen)
}


# Examples

# Laplace Density vs histogram
tempf = Vectorize(function(x) dlap(x,10,2))
set.seed(111)
hist(rlap(10000,10,2),breaks=250,probability=T,ylim=c(0,0.3),xlab="x",ylab="Density",main="Simulated data")
curve(tempf,col="red",lwd=2,add=T,n=1000)
box()

# Evaluation
qlap(0.3)
plap(-0.5108256)
dlap(0.3)

# two-piece Laplace density vs histogram
set.seed(111)
hist(rtp3(10000,0,1,0.5,param="eps",FUN=rlap),breaks=250,probability=T,ylim=c(0,0.55),xlab="x",ylab="Density",main="Simulated data")
dtplap = Vectorize(function(x) dtp3(x,0,1,0.5,FUN=dlap,param="eps"))
curve(dtplap,-15,5,n=1000,lwd=2,xlab="x",ylab="Density",add=T,col="red")
box()
```
