---
title: "The Hyperbolic Secant Distribution"
author: '[F. Javier Rubio](https://sites.google.com/site/fjavierrubio67/)'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Hyperbolic Secant Distribution

The [Hyperbolic Secant distribution](https://en.wikipedia.org/wiki/Hyperbolic_secant_distribution) is a distribution with support on ${\mathbb R}$. It is a continuous, symmetric, and tractable distribution with heavier tails than those of the Normal distribution. The probability density function of the Hyperbolic Secant distribution is

$$f(x; \mu, \sigma )={\frac{1}{2\sigma}}\;\operatorname {sech}\!\left({\frac{\pi(x-\mu) }{2\sigma}}\right)\!, $$
where $\mu \in {\mathbb R}$, $\sigma>0$ and
$${\displaystyle \operatorname {sech} (x)={\frac {1}{\cosh (x)}}={\frac {2}{e^{x}+e^{-x}}}={\frac {2e^{x}}{e^{2x}+1}}}.$$
The cumulative distribution function of the Hyperbolic Secant distribution is
$$F(x; \mu, \sigma ) = {\frac  {2}{\pi }}\arctan \!\left[\exp \left(\frac{\pi(x-\mu) }{2\sigma}\right)\right]\!. $$
The mean of the hyperbolic secant distribution is $E[X] = \mu$, and the variance is $Var[X]=\sigma^2$. Asymmetric versions of this distribution using the [two piece construction](http://rpubs.com/FJRubio/TPD) are discussed in @steel:2015.

The following R code shows the implementation of the pdf, the cdf, the quantile function, random number generation, and moments associated to the Hyperbolic Secant distribution.

```{r}
#########################################################################################
# Parameters
#########################################################################################
# mu in R
# sigma > 0
#########################################################################################

#------------------------------------------------------------------
# PDF
#------------------------------------------------------------------

dsech <- Vectorize(function(x,mu,sigma,log = FALSE){
  logden <-  -log(2) - log(sigma) - log( cosh( 0.5*pi*(x-mu)/sigma ) ) 
  val <- ifelse(log, logden, exp(logden)) 
  return(val)
})

#------------------------------------------------------------------
# CDF
#------------------------------------------------------------------

psech <- Vectorize(function(x,mu,sigma,log.p = FALSE){
  logcdf <-  log(2) - log(pi) + log( atan( exp( 0.5*pi*(x-mu)/sigma ) ) )
  val <- ifelse(log.p, logcdf, exp(logcdf))
  return(val)
})

#------------------------------------------------------------------
# Quantile function
#------------------------------------------------------------------

qsech <- Vectorize(function(p,mu,sigma){
  val <- sigma*2*log( tan( 0.5*pi*p ) )/pi + mu
  return(val)
})

#------------------------------------------------------------------
# Random number generation
#------------------------------------------------------------------

rsech <- function(n,mu,sigma){
  u <- runif(n)
  val <-  sigma*2*log( tan( 0.5*pi*u ) )/pi + mu
  return(val)
}


##############################################
# Examples
##############################################
#--------------------
mu0 <- 0
sigma0 <- 1
#--------------------
# PDF
sim <- rsech(n=10000, mu0, sigma0)
tempf <- Vectorize(function(x) dsech(x,mu0, sigma0))
hist(sim, breaks = 50, probability = T, xlab= "x", ylab = "Density", cex.axis = 1.5, cex.lab = 1.5, main = "")
curve(tempf,-6,6,n=1000, add= T, lwd = 2)
box()

# CDF
sim <- rsech(n=10000, mu0, sigma0)
tempf <- Vectorize(function(x) psech(x,mu0, sigma0))
plot(ecdf(sim), xlab= "x", ylab = "CDF", main = "", cex.axis = 1.5, cex.lab = 1.5)
curve(tempf,-6,6,n=1000, add= T, lwd = 2, col="red", lty = 2)
box()

# Sample Mean
mean(sim)
# Sample Variance
var(sim)

#--------------------
mu0 <- 1
sigma0 <- 2
#--------------------
# PDF
sim <- rsech(n=10000, mu0, sigma0)
tempf <- Vectorize(function(x) dsech(x,mu0, sigma0))
hist(sim, breaks = 50, probability = T, xlab= "x", ylab = "Density", cex.axis = 1.5, cex.lab = 1.5, ylim = c(0,0.25), main = "")
curve(tempf,-12,12,n=1000, add= T, lwd = 2)
box()

# CDF
sim <- rsech(n=10000, mu0, sigma0)
tempf <- Vectorize(function(x) psech(x,mu0, sigma0))
plot(ecdf(sim), xlab= "x", ylab = "CDF", main = "", cex.axis = 1.5, cex.lab = 1.5)
curve(tempf,-12,12,n=1000, add= T, lwd = 2, col="red", lty = 2)
box()

# Sample Mean
mean(sim)
# Sample Variance
var(sim)

#--------------------
mu0 <- -1
sigma0 <- 0.5
#--------------------
# PDF
sim <- rsech(n=10000, mu0, sigma0)
tempf <- Vectorize(function(x) dsech(x,mu0, sigma0))
hist(sim, breaks = 50, probability = T, xlab= "x", ylab = "Density", cex.axis = 1.5, cex.lab = 1.5, main = "")
curve(tempf,-5,5,n=1000, add= T, lwd = 2)
box()

# CDF
sim <- rsech(n=10000, mu0, sigma0)
tempf <- Vectorize(function(x) psech(x,mu0, sigma0))
plot(ecdf(sim), xlab= "x", ylab = "CDF", main = "", cex.axis = 1.5, cex.lab = 1.5)
curve(tempf,-5,5,n=1000, add= T, lwd = 2, col="red", lty = 2)
box()

# Sample Mean
mean(sim)
# Sample Variance
var(sim)
```

# References