---
title: "The Power Generalised Weibull Distribution"
author: '[F. Javier Rubio](https://sites.google.com/site/fjavierrubio67/)'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: true
    toc_float: true
bibliography: references.bib
---

# The Power Generalised Weibull Distribution
The Power Generalised Weibull (PGW) distribution [@nikulin:2009] is a three-parameter distribution with support on ${\mathbb R}_+$. The corresponding hazard function can accomodate bathtub, unimodal and monotone (increasing and decreasing) hazard shapes. The PGW distribution has become popular in survival analysis given the tractability of its hazard and survival functions. Other flexible distributions that can account for these hazard shapes are discussed in @rubio:2019 and @jones:2015. See also [The Exponentiated Weibull distribution](http://rpubs.com/FJRubio/EWD) and [Simulating survival times from a General Hazard structure with a flexible baseline hazard](http://www.rpubs.com/FJRubio/GHSim).

### Probability Density Function
The pdf of the PGW distribution is
$$f(t;\sigma,\nu,\gamma) = \dfrac{\nu}{\gamma \sigma^\nu}t^{\nu-1} \left[ 1 + \left(\dfrac{t}{\sigma}\right)^\nu\right]^{\left(\frac{1}{\gamma}-1\right)} \exp\left\{ 1- \left[ 1 + \left(\dfrac{t}{\sigma}\right)^\nu\right]^{\frac{1}{\gamma}}
\right\},$$
where $\sigma>0$ is a scale parameter, and $\nu,\gamma >0$ are shape parameters.

### Survival Function
The survival function of the PGW distribution is
$$S(t;\sigma,\nu,\gamma) = \exp\left\{ 1- \left[ 1 + \left(\dfrac{t}{\sigma}\right)^\nu\right]^{\frac{1}{\gamma}}
\right\}.$$

###  Hazard Function
The hazard function of the PGW distribution is
$$h(t;\sigma,\nu,\gamma) = \dfrac{\nu}{\gamma \sigma^\nu}t^{\nu-1} \left[ 1 + \left(\dfrac{t}{\sigma}\right)^\nu\right]^{\left(\frac{1}{\gamma}-1\right)}.$$
The cdf can be obtained as $F(t;\sigma,\nu,\gamma)=1-S(t;\sigma,\nu,\gamma)$, and the cumulative hazard function as $H(t;\sigma,\nu,\gamma) = -\log S(t;\sigma,\nu,\gamma)$, as usual.

### Quantile Function
The quantile function of the PGW distribution is
$$Q(p;\sigma,\nu,\gamma) = \sigma \left[ \left( 1 - \log(1-p) \right)^{\gamma} - 1 \right]^{\frac{1}{\nu}},$$
where $p\in(0,1)$.

The following R code shows the implementation of the pdf, survival function, hazard function, cumulative hazard function, quantile function, and random number generation associated to the PGW distribution. Some illustrative examples are also presented.

# R functions

```{r }
# sigma   : scale parameter
# nu      : shape parameter
# gamma   : shape parameter
# t       : positive argument   
# p       : probability (0,1)
# n       : number of simulations

# Probability Density Function
dpgw <- function(t, sigma, nu, gamma, log = FALSE){
val <- log(nu) - log(gamma) - nu*log(sigma) + (nu-1)*log(t) + 
       (1/gamma - 1)*log( 1 + (t/sigma)^nu ) + 
       ( 1 - ( 1 + (t/sigma)^nu )^(1/gamma) )
  if(log) return(val) else return(exp(val))
}

# PGW Survival Function
spgw <- function(t, sigma, nu, gamma, log.p = FALSE){
  val <- 1 - ( 1 + (t/sigma)^nu )^(1/gamma)
  if(log.p) return(val) else return(exp(val))
}

# PGW Hazard Function
hpgw <- function(t, sigma, nu, gamma, log = FALSE){
  val <- log(nu) - log(gamma) - nu*log(sigma) + (nu-1)*log(t) + 
    (1/gamma - 1)*log( 1 + (t/sigma)^nu )
  if(log) return(val) else return(exp(val))
}

# PGW Cumulative Hazard Function
chpgw <- function(t, sigma, nu, gamma){
  val <- -1 + ( 1 + (t/sigma)^nu )^(1/gamma)
  return(val) 
}

# Quantile Function
qpgw <- function(p, sigma, nu, gamma){
  out <- sigma*(  ( 1 - log(1-p) )^gamma - 1 )^(1/nu)
  return(out)
}

# Random Number Generation Function
rpgw <- function(n, sigma, nu, gamma){
  p <- runif(n)
  out <- sigma*(  ( 1 - log(1-p) )^gamma - 1 )^(1/nu)
  return(as.vector(out))
}
```


# Examples

### PDF and Random number generation
```{r }
# Simulated data
sigma0 = 1; nu0 = 3; gamma0 = 2
set.seed(123)
sim <- rpgw(n = 1000, sigma = sigma0, nu = nu0, gamma = gamma0)
# True pdf
true.pdf <- Vectorize(function(t) dpgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
hist(sim, probability = TRUE, cex.axis = 1.5, cex.lab = 1.5, breaks = 25, main = "", xlab = "")
curve(true.pdf, 0, max(sim), lwd = 2, col = "blue", add = T)
box()

sigma0 = 1; nu0 = 0.75; gamma0 = 0.75
set.seed(123)
sim <- rpgw(n = 1000, sigma = sigma0, nu = nu0, gamma = gamma0)
# True pdf
true.pdf <- Vectorize(function(t) dpgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
hist(sim, probability = TRUE, cex.axis = 1.5, cex.lab = 1.5, breaks = 25, main = "", xlab = "")
curve(true.pdf, 0, max(sim), lwd = 2, col = "blue", add = T)
box()

sigma0 = 1; nu0 = 1.5; gamma0 = 0.75
set.seed(123)
sim <- rpgw(n = 1000, sigma = sigma0, nu = nu0, gamma = gamma0)
# True pdf
true.pdf <- Vectorize(function(t) dpgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
hist(sim, probability = TRUE, cex.axis = 1.5, cex.lab = 1.5, breaks = 25, main = "", xlab = "")
curve(true.pdf, 0, max(sim), lwd = 2, col = "blue", add = T)
box()
```

### Some Hazard Function shapes

```{r}
sigma0 = 0.75; nu0 = 1.5; gamma0 = 3
haz <- Vectorize(function(t) hpgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
curve(haz,0,10, lwd = 2, col = "red", xlab = "", main = "", ylab = "Hazard Function", cex.axis = 1.5, cex.lab = 1.5, n = 250)

sigma0 = 10; nu0 = 2; gamma0 = 0.5
haz <- Vectorize(function(t) hpgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
curve(haz,0,10, lwd = 2, col = "red", xlab = "", main = "", ylab = "Hazard Function", cex.axis = 1.5, cex.lab = 1.5, n = 250)

sigma0 = 5; nu0 = 2; gamma0 = 1.5
haz <- Vectorize(function(t) hpgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
curve(haz,0,10, lwd = 2, col = "red", xlab = "", main = "", ylab = "Hazard Function", cex.axis = 1.5, cex.lab = 1.5, n = 250)

sigma0 = 1; nu0 = 0.75; gamma0 = 0.75
haz <- Vectorize(function(t) hpgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
curve(haz,0,10, lwd = 2, col = "red", xlab = "", main = "", ylab = "Hazard Function", cex.axis = 1.5, cex.lab = 1.5, n = 250)
```

### CDF and Random Number Generation (vs ECDF)

```{r}
# Simulated data
sigma0 = 1; nu0 = 3; gamma0 = 2
set.seed(123)
sim <- rpgw(n = 1000, sigma = sigma0, nu = nu0, gamma = gamma0)
# True cdf
true.cdf <- Vectorize(function(t) 1-spgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
plot(ecdf(sim), cex.axis = 1.5, cex.lab = 1.5, main = "", xlab = "", ylab = "CDF vs ECDF")
curve(true.cdf, 0, max(sim), lwd = 2, col = "gray", add = T)
box()

sigma0 = 1; nu0 = 0.75; gamma0 = 0.75
set.seed(123)
sim <- rpgw(n = 1000, sigma = sigma0, nu = nu0, gamma = gamma0)
# True cdf
true.cdf <- Vectorize(function(t) 1-spgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
plot(ecdf(sim), cex.axis = 1.5, cex.lab = 1.5, main = "", xlab = "", ylab = "CDF vs ECDF")
curve(true.cdf, 0, max(sim), lwd = 2, col = "gray", add = T)
box()

sigma0 = 1; nu0 = 1.5; gamma0 = 0.75
set.seed(123)
sim <- rpgw(n = 1000, sigma = sigma0, nu = nu0, gamma = gamma0)
# True cdf
true.cdf <- Vectorize(function(t) 1-spgw(t, sigma = sigma0, nu = nu0, gamma = gamma0))
plot(ecdf(sim), cex.axis = 1.5, cex.lab = 1.5, main = "", xlab = "", ylab = "CDF vs ECDF")
curve(true.cdf, 0, max(sim), lwd = 2, col = "gray", add = T)
box()
```

# References