---
title: "The Kumaraswamy Distribution"
author: '[F. Javier Rubio](https://sites.google.com/site/fjavierrubio67/)'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Kumaraswamy Distribution

The [Kumaraswamy distribution](https://en.wikipedia.org/wiki/Kumaraswamy_distribution) is a distribution with support on $[0,1]$. It represents an attractive alternative to the [Beta distribution](https://en.wikipedia.org/wiki/Beta_distribution) as its pdf and cdf can be written in closed form. An extensive comparison of these two distribution can be found at [@jones:2009]. The probability density function of the Kumaraswamy distribution is

$${\displaystyle f(x;a,b)=abx^{a-1}{(1-x^{a})}^{b-1},\ \ {\mbox{where}}\ \ x\in [0,1].} $$
where $a>0$, $b>0$. The cumulative distribution function of the Kumaraswamy distribution is
$${\displaystyle F(x;a,b)={\displaystyle 1-(1-x^{a})^{b}},\ \ {\mbox{where}}\ \ x\in [0,1].} $$
Estimation of the parameters of this distribution using the method of moments can be found at [this RPbus](http://rpubs.com/FJRubio/KumaMM).

The following R code shows the implementation of the pdf, the cdf, the quantile function, random number generation, and moments associated to the Kumaraswamy distribution.

```{r}
#########################################################################################
# Parameters
#########################################################################################
# a > 0
# b > 0
# 0 < x < 1
#########################################################################################

#------------------------------------------------------------------
# PDF
#------------------------------------------------------------------

dkuma <- Vectorize(function(x,a,b,log = FALSE){
logden <-  log(a) + log(b) + (a-1)*log(x) + (b-1)*log(1-x^a)
val <- ifelse(log, logden, exp(logden)) 
return(val)
})

#------------------------------------------------------------------
# CDF
#------------------------------------------------------------------

pkuma <- Vectorize(function(q,a,b,log.p = FALSE){
  cdf <-  1 - (1-q^a)^b
  val <- ifelse(log.p, log(cdf), cdf)
  return(val)
})

#------------------------------------------------------------------
# Quantile function
#------------------------------------------------------------------

qkuma <- Vectorize(function(u,a,b){
  val <-  ( 1 - (1-u)^(1/b) )^(1/a)
  return(val)
})

#------------------------------------------------------------------
# Random number generation
#------------------------------------------------------------------

rkuma <- function(n,a,b){
  u <- runif(n)
  val <-  ( 1 - (1-u)^(1/b) )^(1/a)
  return(val)
}


#------------------------------------------------------------------
# Moments of order n of the Kumaraswamy(a,b) distribution
#------------------------------------------------------------------
mn <- function(a,b,n){
  log.num <- log(b) + lgamma(1 + n/a) + lgamma(b)
  log.den <- lgamma(1 + b + n/a)
  return(exp(log.num-log.den))
}



##############################################
# Examples
##############################################
a0 <- 1
b0 <- 1

sim <- rkuma(n=10000, a0, b0)
tempf <- Vectorize(function(x) dkuma(x,a0,b0))
hist(sim, breaks = 50, probability = T)
curve(tempf,0,1,n=1000, add= T, lwd = 2)
box()


#----------------------
a0 <- 2
b0 <- 3
#----------------------
sim <- rkuma(n=10000, a0, b0)
tempf <- Vectorize(function(x) dkuma(x,a0,b0))
hist(sim, breaks = 50, probability = T)
curve(tempf,0,1,n=1000, add= T, lwd = 2)
box()


#----------------------
a0 <- 3
b0 <- 2
#----------------------
sim <- rkuma(n=10000, a0, b0)
tempf <- Vectorize(function(x) dkuma(x,a0,b0))
hist(sim, breaks = 50, probability = T)
curve(tempf,0,1,n=1000, add= T, lwd = 2)
box()



#----------------------
a0 <- 20
b0 <- 30
#----------------------
sim <- rkuma(n=10000, a0, b0)
tempf <- Vectorize(function(x) dkuma(x,a0,b0))
hist(sim, breaks = 50, probability = T)
curve(tempf,0,1,n=1000, add= T, lwd = 2)
box()


#----------------------
a0 <- 30
b0 <- 20
#----------------------
sim <- rkuma(n=10000, a0, b0)
tempf <- Vectorize(function(x) dkuma(x,a0,b0))
hist(sim, breaks = 50, probability = T)
curve(tempf,0,1,n=1000, add= T, lwd = 2)
box()



#----------------------
a0 <- 2
b0 <- 3
#----------------------
sim <- rkuma(n=10000, a0, b0)
tempf <- Vectorize(function(x) dkuma(x,a0,b0))
hist(sim, breaks = 50, probability = T)
curve(tempf,0,1,n=1000, add= T, lwd = 2)
box()


#----------------------
a0 <- 2.2
b0 <- 3
#----------------------
sim <- rkuma(n=10000, a0, b0)
tempf <- Vectorize(function(x) dkuma(x,a0,b0))
hist(sim, breaks = 50, probability = T)
curve(tempf,0,1,n=1000, add= T, lwd = 2)
box()
```

# References